---
title: "RWorksheet#4c_Parrenas"
author: "Erikka Jane Parrenas"
date: "2023-11-22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#1. Use the dataset mpg
#a. Show your solutions on how to import a csv file into the environment
library(readr)
mpg <- read.csv("/cloud/project/mpg.csv")
mpg
```
```{r}
#b. Which variables from mpg dataset are categorical?
#The categorical variables in the mpg dataset are:
data(mpg)
str(mpg)

```

#c. Which are continuous variables?
#The continuous variables in the mpg dataset are:

#cty: The city mpg (miles per gallon) of the car.
#hwy: The highway mpg (miles per gallon) of the car.
#displ: The engine displacement (volume) of the car.

```{r}
#2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.

manufacturers <- table(mpg$manufacturer)
maxModels <- which.max(manufacturers)
manufacturer_with_most_models <- names(manufacturers)[maxModels]

print(paste("The manufacturer with the most models is:", manufacturer_with_most_models))

models <- table(mpg$model)
maxVariations <- which.max(models)
model_with_most_variations <- names(models)[maxVariations]

print(paste("The model with the most variations is:", model_with_most_variations))
```
```{r}
#2.1a. Group the manufacturers and find the unique models. Show your codes and result.
library(dplyr)
manuModels <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_models = n_distinct(model))
print(manuModels)
```

```{r}
##2.1b. Graph the result by using plot() and ggplot(). Write the codes and its result.
library(ggplot2)

plot(ggplot(manuModels, aes(x = reorder(manufacturer, -unique_models), y = unique_models, fill = unique_models)) +
 geom_bar(stat = "identity", width = 0.8) + 
 labs(title = "Number of Manufacturer's Unique Models", 
      x = "Manufacturer",
      y = "Number of Unique Models") +
   
 theme_minimal() +
 scale_fill_gradient(low = "brown", high = "green") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))
       
```

```{r}
##2.2a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
#With the help of the ggplot code, an mpg data plot is created, with each point denoting a different manufacturer's car model.

#An ggplot object, which is a specification of the data and a mapping of the data to the visual objects, is created using the ggplot() method. The aesthetic mapping is defined by the aes() function. In this instance, it associates the x and y coordinates with the model and manufacturer variables, respectively. A point layer can be added to the plot using the geom_point() method. It will by default plot the points in a scatter plot. The storyline is finally shown.

#2b. For you, is it useful? If not, how could you modify the data to make it more informative?

#yes, it is useful. This code is helpful for understanding the relationship between car models and manufacturers in the mpg data. In order to assist the viewer and prevent misunderstanding, I will add color to the points dependent on various factors. I will also modify the size of the variable titles according to their angle to make them easier to read.
```

```{r}
##3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
library(ggplot2)
library(dplyr)

data(mpg)


mean_displ_df <- mpg %>%
 group_by(year, model) %>%
 summarise(mean_displ = mean(displ)) %>%
 arrange(desc(mean_displ)) %>%
 filter(row_number() < 20) 

plot <- ggplot(mean_displ_df, aes(x = year, y = mean_displ, fill = model)) +
 geom_bar(stat = "identity", position = "dodge") +
 theme_minimal() +
 guides(fill = guide_legend(ncol = 2)) 
 labs(title = "Average Engine Displacement over the years for the top 20 models",
       x = "Year",
       y = "Engine Displacement",
       fill = "Model")

print(plot)
```

```{r}
#4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

library(dplyr)
data(mpg)


countpermodel <- mpg %>%
  group_by(model) %>%
  summarise(num_cars = n())


print(countpermodel)
```

```{r}
#a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.

library(ggplot2)
library(dplyr)

data(mpg)

dataSummary <- mpg %>%
  count(model) %>%
  arrange(desc(n)) %>%
  slice(1:20)

top_models <- dataSummary$model
palette <- scales::hue_pal()(length(top_models))

dataSummary <- dataSummary %>%
  mutate(color = palette[match(model, top_models)])

ggplot(dataSummary, aes(x = reorder(model, n), y = n, fill = model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Top 20 Car Models by Count",
    x = "Car Models",
    y = "Total Cars"
  ) +
  scale_fill_manual(values = palette, name = "Car Models", breaks = dataSummary$model) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.key.size = unit(0.1, "cm"),
    plot.title = element_text(hjust = 0.5)  
  )

```

```{r}
#b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

library(ggplot2)
library(dplyr)

data(mpg)

dataSummary <- mpg %>%
  count(model) %>%
  arrange(desc(n)) %>%
  slice(1:20)


top_models <- dataSummary$model
palette <- scales::hue_pal()(length(top_models))

dataSummary <- dataSummary %>%
  mutate(color = palette[match(model, top_models)])

ggplot(dataSummary, aes(x = reorder(model, n), y = n, fill = model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Top 20 Car Models by Count",
    y = "Car Models",  
    x = "Total Cars"
  ) +
  scale_fill_manual(values = palette, name = "Car Models", breaks = dataSummary$model) +
  coord_flip() + 
  theme_minimal() +
  theme(
    legend.key.size = unit(0.1, "cm"),
    plot.title = element_text(hjust = 0.5)  
  )

```

```{r}
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(
    title = "Relationship between No. of Cylinders and Engine Displacement",
    x = "Number of Cylinders",
    y = "Engine Displacement"
  )
```

#a. How would you describe its relationship? Show the codes and its result.

#By using line regression to illustrate the relationship between the number of cylinders and displacement, it can be seen that an increase in cylinder count corresponds with an increase in engine size.


```{r}
library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +  
  labs(
    title = "Relationship between No. of Cylinders and Engine Displacement",
    x = "Number of Cylinders",
    y = "Engine Displacement"
  )

```


#6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?

#Engine displacement (displ) is shown against highway miles per gallon (hwy) in a scatter plot, with a continuous variable represented across the points by the color gradient of city miles per gallon (cty).

#Instead of showing a clear correlation between engine displacement and highway and city miles per gallon (displ and hwy), the color gradient based on cty represents the difference in city MPG across the scatter plot.

```{r}
library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(
    title = "Relationship between Engine Displacement and Highway MPG",
    x = "Engine Displacement",
    y = "Highway MPG"
  ) +
  scale_color_gradient(low = "black", high = "magenta")  

```

```{r}
#6. Import the traffic.csv onto your R environment.
library(readr)
traffic <- read.csv("/cloud/project/traffic.csv")
traffic
```

```{r}
#a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.

observations <- nrow(traffic)
variables <- names(traffic)

cat("Number of observations:", observations, "\n")
cat("The variables are:", variables, "\n")
```

```{r}
#b. subset the traffic dataset into junctions. What is the R codes and its output?

junctions1 <- subset(traffic, Junction == 1)
junctions2 <- subset(traffic, Junction == 2)
junctions3 <- subset(traffic, Junction == 3)
junctions4 <- subset(traffic, Junction == 4) 

#The output are:
junctions1  
junctions2
junctions3
junctions4
```

```{r}
#c. Plot each junction in a using geom_line(). Show your solution and output.

ggplot(junctions1, aes(x = DateTime, y = Vehicles, color = "Junction 1")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 1",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
 theme(plot.title = element_text(hjust = 0.5))

#Junction 2
ggplot(junctions2, aes(x = DateTime, y = Vehicles, color = "Junction 2")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 2",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
  theme(plot.title = element_text(hjust = 0.5))

#Junction 3
ggplot(junctions3, aes(x = DateTime, y = Vehicles, color = "Junction 3")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 3",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
 theme(plot.title = element_text(hjust = 0.5))

#Junction 4
ggplot(junctions4, aes(x = DateTime, y = Vehicles, color = "Junction 4")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 4",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#7. From alexa_file.xlsx, import it to your environment
library(readxl)
alexa_file <- read_excel("/cloud/project/RWorksheet#4a&4b/RWorksheet#4b/alexa_file.xlsx")
head(alexa_file)
```

```{r}
#a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.

observations <- nrow(alexa_file)
columns <- ncol(alexa_file)

cat("Number of observations:", observations, "\n")
cat("Number of columns:", columns, "\n")

```

```{r}
#b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.

library(dplyr)

result <- alexa_file %>%
  group_by(variation) %>%
  summarise(total_variations = n()) 

print(result)
```

```{r}
#c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.

library(ggplot2)

variat <- ggplot(result, aes(x = variation, y = total_variations, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Variations Count",
       x = "Variation",
       y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_discrete(name = "Variation") +
  guides(fill = guide_legend(ncol = 2)) +
  theme(plot.title = element_text(hjust = 0.5)) 

print(variat)

#The variation of the Alexa file are shown below, with the sum of each variation as well as each variation's name and color indicated to help the viewer evaluate the figure.  Compared to the others, the Black Dot variety is either more well-known or appears much more frequently. The caption, which is divided into two columns, makes it simple to understand which hue corresponds to each kind of variation.

```

```{r}
#d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.

library(dplyr)
library(ggplot2)

alexa_file$date <- as.Date(alexa_file$date)
alexa_file$month <- format(alexa_file$date, "%m")

monthcount <- alexa_file %>%
   count(month)

 ploto <- ggplot(monthcount, aes(x = as.integer(month), y = n, color = "Reviews")) +
  geom_line(size = 1) +
  labs(title = "Number of Verified Reviews Over Time",
       x = "Month",
       y = "Number of Verified Reviews",
       color = "Legend Title") +  # Change legend title
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_color_manual(values = c("red"), labels = c("Reviews")) +  
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

print(ploto)
```

```{r}
#e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.

library(dplyr)
library(ggplot2)

vari_ratings <- alexa_file %>%
  group_by(variation) %>%
  summarize(avg_rating = mean(rating))
print(vari_ratings)

highest <- vari_ratings %>% 
  filter(avg_rating == max(avg_rating))
print(highest) 

vari_names <- vari_ratings$variation
average_ratings <- vari_ratings$avg_rating

barplot(average_ratings, names.arg = vari_names, col = "blue",
        main = "Average Rating by Variation",
        xlab = "Variation", ylab = "Average Rating",
        cex.axis = 0.8, cex.names = 0.5, las = 2)  

top_variation <- vari_names[which.max(average_ratings)]
top_rating <- max(average_ratings)

cat("The variation with the highest average rating is:", top_variation, "with an average rating of", top_rating)
```

